Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-02-01 paper/main.tex

Designing and building procedures and tools for mapping out knowledge touches
on deep questions about what it means to learn. For example, how do we acquire
conceptual knowledge? Memorizing course lectures or textbook chapters by rote
can lead to the superficial \textit{appearance} of understanding the underlying
content, but achieving true conceptual understanding seems to require something
deeper and richer. Does conceptual understanding entail connecting newly
acquired information to the scaffolding of one's existing knowledge or
experience~\citep{BlayEtal06,CaraMaho03, ConsEtal16, DeacEtal04, SimoEtal04}?
Or weaving a lecture's atomic elements (e.g., its component words) into a
structured network that describes how those individual elements are related?
Conceptual understanding could also involve building a mental model that
transcends the meanings of those individual atomic elements by reflecting the
deeper meaning underlying the gestalt whole~\citep{Kint70, Macl05, ScotEtal07}.

@@

Our primary research goal is to advance our understanding of what it means to
acquire deep, real-world conceptual knowledge. Traditional laboratory
approaches to studying learning and memory (e.g., list-learning studies) often
draw little distinction between memorization and understanding. Instead, these
studies typically focus on whether information is effectively encoded or
retrieved, rather than whether the information is \textit{understood}.
Approaches to studying conceptual learning, such as category learning
experiments, can begin to investigate the distinction between memorization and
understanding, often by training participants to distinguish arbitrary or
random features in otherwise meaningless categorized stimuli. However the
objective of real-world training, or learning from life experiences more
generally, is often to develop new knowledge that may be applied in
\textit{useful} ways in the future. In this sense, the gap between modern
learning theories and modern pedagogical approaches that inform classroom learning
strategies is enormous: most of our theories about \textit{how} people learn
are inspired by experimental paradigms and models that have only peripheral
relevance to the kinds of learning that students and teachers actually
seek~\citep{Macl05, HallGree08}. To help bridge this gap, our study uses course
materials from real online courses to inform, fit, and test models of
real-world conceptual learning. We also provide a demonstration of how our
models can be used to construct ``maps'' of what students know, and how their
knowledge changes with training. In addition to helping to visualize knowledge
(and changes in knowledge), we hope that such maps might lead to real-world
tools for improving how we educate.

>>>>

Designing and building procedures and tools for mapping out knowledge touches
on deep questions about what it means to learn. For example, how do we acquire
conceptual knowledge? Memorizing course lectures or textbook chapters by rote
can lead to the superficial \textit{appearance} of understanding the underlying
content, but achieving true conceptual understanding seems to require something
deeper and richer. Does conceptual understanding entail connecting newly
acquired information to the scaffolding of one's existing knowledge or
experience~\citep{BlayEtal06,CaraMaho03, ConsEtal16, DeacEtal04, SimoEtal04}?
Or weaving a lecture's atomic elements (e.g., its component words) into a
structured network that describes how those individual elements are
related~\citep{LeeChen22}? Conceptual understanding could also involve building
a mental model that transcends the meanings of those individual atomic elements
by reflecting the deeper meaning underlying the gestalt whole~\citep{Kint70,
Macl05, ScotEtal07}.

@@

Our primary research goal is to advance our understanding of what it means to
acquire deep, real-world conceptual knowledge. Traditional laboratory
approaches to studying learning and memory (e.g., list-learning studies) often
draw little distinction between memorization and understanding. Instead, these
studies typically focus on whether information is effectively encoded or
retrieved, rather than whether the information is \textit{understood}.
Approaches to studying conceptual learning, such as category learning
experiments, can begin to investigate the distinction between memorization and
understanding, often by training participants to distinguish arbitrary or
random features in otherwise meaningless categorized stimuli. However the
objective of real-world training, or learning from life experiences more
generally, is often to develop new knowledge that may be applied in
\textit{useful} ways in the future. In this sense, the gap between modern
learning theories and modern pedagogical approaches that inform classroom
learning strategies is enormous: most of our theories about \textit{how} people
learn are inspired by experimental paradigms and models that have only
peripheral relevance to the kinds of learning that students and teachers
actually seek~\citep{Macl05, HallGree08}. To help bridge this gap, our study
uses course materials from real online courses to inform, fit, and test models
of real-world conceptual learning. We also provide a demonstration of how our
models can be used to construct ``maps'' of what students know, and how their
knowledge changes with training. In addition to helping to visualize knowledge
(and changes in knowledge), we hope that such maps might lead to real-world
tools for improving how we educate. Taken together, our work shows that
existing course materials and evaluative tools like short multiple-choice
quizzes may be leveraged to gain highly detailed insights into what students
know and how they learn.
